Random Forests", OAs

### 2023-11-13 src/la_RF.Rmd
rf

---
title: "Random Forests"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

@@

```{r}
#spatial weight matrix
sf_use_s2(FALSE)
nb <- poly2nb(la, queen = T, snap = 1)
listw <- nb2listw(nb, style = "W", zero.policy = T)

@@

```{r}
#install.packages('SDPDmod')
library(SDPDmod)

@@

```{r}
# London = Islington 
sf_use_s2(FALSE)
help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
  left_join(df, by = c("LAD21CD"="ladcd")) %>% dplyr::select(LAD21CD, year, n)

@@

knn.l <- nb2listw(knn)

@@

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   #seeds = seeds,
                   allowParallel = T,
                   savePredictions = 'final')

@@

for (i in regions){
  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
                      n.slag + n.nearest.city + year, #n.lag +  + n.l.slag
                     data = df %>% filter(RGN21CD==i), trControl = tc,
                     method = "xgbTree", # rf "rpart", # rf lm
                     na.action = na.omit,
                     #preProc = c("center", "scale"),
                     importance = TRUE)
  assign(paste0("model.all",as.character(i)), model.all)
}

@@

# use the model trainned in year t to predict io of year t + 1
#pred.by.year.all <- data.frame(matrix(NA,nrow=336, ncol = 0))
for (i in regions){
  for (j in regions[!regions == i]){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", i, ".on.", j), pred)
    #pred.by.year.all <- rbind(pred.by.year.all, pred)
  }
}

df.predictions <- data.frame()
for (i in regions){
  for (j in regions[!regions == i]){
    d <- paste0("region.predict.model.", i, ".on.", j)
    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
    d <- d[,c(5, 2, 4, 6, 1)]
    colnames(d)[5] <- "predictions"
    d <- d %>% mutate(trained.on = i)
    df.predictions <- rbind(d, df.predictions)
  }
}

@@

# pred.by.year to list by year
pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11

# calculate metrics for every year
rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$n))  # CHANGE n WITH growth
rf.year.all.metrics %>% as.data.frame() %>%
  round(2) %>%
  kable()

# there are 2 columns named year, so I drop one
pred.by.region.all$year <- NULL

ggplot(data = df.predictions, aes(x = n, y = predictions)) +
  geom_point(colour = "blue") +
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
  facet_wrap(~ test.train, ncol = 2) +
  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
  ggtitle("Predicted vs. actual trade by year")
```

>>>>

---
title: "Random Forests, LAs"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

@@

```{r, eval=F}
#spatial weight matrix
sf_use_s2(FALSE)
nb <- poly2nb(la, queen = T, snap = 1)
listw <- nb2listw(nb, style = "W", zero.policy = T)

@@

```{r, eval=F}
#install.packages('SDPDmod')
library(SDPDmod)

@@

```{r}
# London = Islington 
sf_use_s2(FALSE)
path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la.sf <- st_read(path.geo)

help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
  left_join(df, by = c("LAD21CD"="ladcd")) %>% dplyr::select(LAD21CD, year, n)

@@

sf_use_s2(FALSE)
la.c <- st_centroid(la.sf)
kn <- knearneigh(la.c, k = 5)
knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
knn.l <- nb2listw(knn)

@@

@@

for (i in regions){
  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
                      n.slag + n.nearest.city + year, #n.lag +  + n.l.slag
                     data = df %>% filter(RGN21CD==i), trControl = tc,
                     method = "rf", # xgbTree "rpart", # rf lm
                     na.action = na.omit,
                     #preProc = c("center", "scale"),
                     importance = TRUE)
  assign(paste0("model.all",as.character(i)), model.all)
}

@@

# train on one region, test on all the rest

df.predictions <- data.frame()
for (i in regions){
  for (j in regions[!regions == i]){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", i, ".on.", j), pred)

    d <- paste0("region.predict.model.", i, ".on.", j)
    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
    d <- d[,c(5, 2, 4, 6, 1)]
    colnames(d)[5] <- "predictions"
    d <- d %>% mutate(trained.on = i)
    df.predictions <- rbind(d, df.predictions)
  }
}

@@

# split to list by region pair
pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11

# calculate metrics for every region pair 
rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$n))  # CHANGE n WITH growth

path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup) %>% 
  dplyr::select(RGN21CD, RGN21NM) %>% 
  distinct() %>% 
  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))

rf.year.all.metrics %>% 
  as.data.frame() %>%
  add_rownames(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "train.test") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  separate(train.test, c("train", "test"), remove = F) %>% 
  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
  rename(train.region=RGN21NM) %>% 
  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
  rename(test.region=RGN21NM) %>% 
  arrange(Rsquared)
  #summarise(r = range(Rsquared))
  #round(2) %>%
  #kable()
```

### test plots

```{r}
#plots
ggplot(data = df.predictions, aes(x = n, y = predictions)) +
  geom_point(colour = "blue") +
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
  facet_wrap(~ test.train, ncol = 2) +
  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
  ggtitle("Predicted vs. actual trade by year")
```

### 2023-11-13 src/oa_RF.Rmd
rf

---
title: "diffusion, websites with 1 pc"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

@@

library(tidyverse)
library(rprojroot)
library(rgdal)
library(rgeos)
library(spdep)
library(tmap)
library(maptools)
library(knitr)
library(REAT)
library(tidygeocoder)
library(geosphere)
library(broom)
library(plm)
library(lmtest)

@@

# build oa for GB
oa.gb <- spRbind(oa.ew, oa.sc)
rm(oa.ew, oa.sc)

@@

# get GB
path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
gb <- readOGR(path.gb)
# spatial transformations
gb <- spTransform(gb, CRS("+init=epsg:4326"))

@@

>>>>

---
title: "Random Forests", OAs
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

@@

library(tidyverse)
library(rprojroot)
library(rgdal)
library(rgeos)
library(sf)
library(spdep)
#library(tmap)
#library(maptools)
library(knitr)
library(REAT)
library(tidygeocoder)
library(geosphere)
library(broom)
#library(foreach)
library(doParallel)
library(raster)
library(plm)
#library(lmtest)

@@

# build oa for GB
oa.gb <- maptools::spRbind(oa.ew, oa.sc)
rm(oa.ew, oa.sc)

@@

# get GB
path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
gb <- readOGR(path.gb)
# spatial transformations
gb <- spTransform(gb, CRS("+init=epsg:4326"))
```

## df for RF

```{r}
# London = Islington 
oa.sf <- st_as_sf(oa.gb)

help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
  left_join(df, by = c("LAD21CD"="ladcd")) %>% dplyr::select(LAD21CD, year, n)

help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21NM,LAD21CD) %>% 
  mutate(cities = c("Leicester",
                    "Bristol",
                    "Manchester",
                    "Liverpool",
                    "Sheffield",
                    "Birmingham",
                    "Leeds",
                    "London",
                    "Edinburgh",
                    "Glasgow"))

dist$dist.city.name <- names(dist)[apply(dist[,1:9], MARGIN = 1, FUN = which.min)]
help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
  rename(nearest.city.LAD21CD = LAD21CD) %>% 
  dplyr::select(ladcd, nearest.city.LAD21CD)

sf_use_s2(FALSE)
la.c <- st_centroid(la.sf)
kn <- knearneigh(la.c, k = 5)
knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
knn.l <- nb2listw(knn)

df <- df %>% drop_na() %>% group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(ladcd) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
  left_join(help2, by = c("ladcd" = "ladcd"))

df <- df %>% left_join(df %>% dplyr::select(ladcd,year, n) %>% 
                         rename(n.nearest.city = n), 
                       by = c("nearest.city.LAD21CD" = "ladcd",
                              "year" = "year")) %>% 
  dplyr::select(-c(LAD21NMW, OBJECTID, BNG_E, BNG_N, LONG, LAT, SHAPE_Length)) %>% 
  relocate(LAD21NM, .after = ladcd)

path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup) %>% dplyr::select(LAD21CD, RGN21CD)

df <- df %>% left_join(lookup.region, by = c("ladcd" = "LAD21CD")) %>% 
  relocate(RGN21CD, .after=year)

# North/South
#lookup.region %>% distinct(RGN21NM, .keep_all = T)
south <- c("E12000007", "E12000008", "E12000009" )
df <- df %>% mutate(south = ifelse(RGN21CD %in% south, 1, 0))

# For growth:
#df <- df %>% mutate(growth = log(n/n.lag))

@@

sf_use_s2(FALSE)
la.c <- st_centroid(la.sf)
kn <- knearneigh(la.c, k = 5)
knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
knn.l <- nb2listw(knn)
